---
title: "Analyzing My Facebook Messages"
author: "Lathan Liou"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(rvest)
library(lubridate)
library(ggplot2)
library(stringr)
```

First, I'm going to analyze the messages for one of my friends. I downloaded the data in HTML format, and I scraped using a CSS selector for date only, not including the user or the message!

```{r}
friend <- read_html("facebook-lathanliou/messages/AkshaySwaminathan_48e865ce79/message.html")
 
friend_df <- friend %>%
  html_nodes("._3-94") %>% 
  html_text() %>%
  data.frame()

#clean data
friend_df <- friend_df %>% 
  rename(moment = ".")

#want to split up into more columns
friend_df$moment <- gsub(" ", ",", friend_df$moment)

#remove extra comma
TrimMult <- function(x, char=" ") {
  return(gsub(paste0("^", char, "*|(?<=", char, ")", char, "|", char, "*$"),
              "", x, perl=T))
}

friend_df$moment <- TrimMult(friend_df$moment, ",")

friend_df <- friend_df %>%
  separate(moment, c("month", "day", "year", "time"), ",")

#group by year
friend_df %>% 
  group_by(year) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = year))
```

Now I want to take a closer look at message frequency, grouping by month over the years.

```{r}
#creating day month year variable
friend_df <- friend_df %>% 
  mutate(dmy = paste(day, month, year, sep = " "))

# reformat dmy variable
friend_df$dmy <- dmy(friend_df$dmy)

#automatically categorizes dmy to weeks
friend_df$week_cut <- as.Date(cut(friend_df$dmy, breaks = "week"))

#automatically categorizes dmy to months
friend_df$month_cut <- as.Date(cut(friend_df$dmy, breaks = "month"))

#monthly messages over time
friend_df %>% 
  group_by(week_cut) %>% 
  summarise(n = n()) %>% 
  arrange(week_cut) %>% 
  ggplot(aes(x = week_cut, y = n)) +
  geom_col() +
  scale_x_date(date_breaks = "1 year") +
  labs(title = "Facebook Messages Exchanged Over Time",
       subtitle = "With One Friend Since 2015",
       x = "Date",
       y = "Number of FB Messages Exchanged") +
  theme(axis.text.x = element_text(angle = 70, hjust = 1),
        panel.background = element_rect(fill = "lightblue"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

The peak corresponds to when I was studying abroad and decided to use Messenger (normally I used Whatsapp). Otherwise the frequency is low because when I'm in the United States, I tend to text rather than use Messenger.

Now, I want to repeat this analysis with my girlfriend.

```{r}
gf<- read_html("facebook-lathanliou/messages/ArdenHu_559c55dd37/message.html")
 
gf_df <- gf %>%
  html_nodes("._3-94") %>% 
  html_text() %>%
  data.frame()

#clean data
gf_df <- gf_df %>% 
  rename(moment = ".")

#want to split up into more columns
gf_df$moment <- gsub(" ", ",", gf_df$moment)

gf_df$moment <- TrimMult(gf_df$moment, ",")

gf_df <- gf_df %>%
  separate(moment, c("month", "day", "year", "time"), ",")

#group by year
gf_df %>% 
  group_by(year) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_col(aes(fill = year))
```

```{r}
#creating day month year variable
gf_df <-gf_df %>% 
  mutate(dmy = paste(day, month, year, sep = " "))

# reformat dmy variable
gf_df$dmy <- dmy(gf_df$dmy)

#automatically categorizes dmy to weeks
gf_df$week_cut <- as.Date(cut(gf_df$dmy, breaks = "week"))

#automatically categorizes dmy to months
gf_df$month_cut <- as.Date(cut(gf_df$dmy, breaks = "month"))

#monthly messages over time
gf_df %>% 
  group_by(week_cut) %>% 
  summarise(n = n()) %>% 
  arrange(week_cut) %>% 
  ggplot(aes(x = week_cut, y = n)) +
  geom_col() +
  scale_x_date(date_breaks = "1 month") +
  labs(title = "Facebook Messages Exchanged Over Time",
       subtitle = "With Bebbin Since 2015",
       x = "Date",
       y = "Number of FB Messages Exchanged") +
  theme(axis.text.x = element_text(angle = 70, hjust = 1),
        panel.background = element_rect(fill = "lightblue"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```
# Sources
\begin{enumerate}

\item http://bradleyboehmke.github.io/2015/12/scraping-html-text.html Learning basic HTML scraping

\item http://www.residualthoughts.com/2018/02/12/analyzing-facebook-messages-in-r/ Inspiration for Facebook Messenger Analysis

\item Max Ghenis https://stackoverflow.com/questions/23274035/removing-multiple-commas-and-trailing-commas-using-gsub?utm_medium=organic&utm_source=google_rich_qa&utm_campaign=google_rich_qa Great function for regex manipulation

\end{enumerate}